Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-07-10 paper_jasa.tex
author description and keywords added

%\documentclass{article}
\documentclass{article}
\usepackage{times}
%\usepackage{natbib}
%\usepackage{multicol}
\RequirePackage{natbib}
\usepackage{amsmath, amssymb, fullpage, amsthm, array, algorithm2e,graphicx,asa}
%\usepackage[dvips]{graphics}

@@

% Article top matter
\title{Validation of Visual Statistical Inference, Applied to Linear Models}
\author{Mahbubul Majumder, Heike Hofmann, Dianne Cook\\
        Department of Statistics, Iowa State University}
%\date{\today}  %\today is replaced with the current date
\maketitle

\begin {abstract}  
Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis. Until recently, there were no formal visual methods in place for determining statistical significance of findings. This changed when \citet{buja:2009} conceptually introduced the lineup protocol for formal tests of visual findings. In this paper this is taken a step further by refining the terminology of visual inference, framing the protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists, and it is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models. A human subjects experiment is conducted using simulated data to provide controlled conditions. Results suggest that the lineup protocol provides results comparable to the conventional tests, out-performs them when data is contaminated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult trials.
\end {abstract}

@@

\hh{Their minimum $p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}$ then follows a Beta distribution with shape parameters 1 and $m-1$, and corresponding distribution function 
\[
F_{p_0} (x) =  1- (1- x) ^{m-1}  \text{ for } x \in [0,1].
\]

\begin{eqnarray*}
P(p_D < p_{0}) &=& 1 - P(p_{0} \le p_D) = 1- \int_0^1  P(p_{0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
&=& 1 - \int_0^1 F_{p_{0}}(t) f_{p_D}(t) dt = 1 - \int_0^1 f_{p_D}(t) dt + \int_0^1 (1-t)^{m-1} f_{p_D}(t) dt =  E\left[ (1 - p_D)^{m-1}\right].
\end{eqnarray*}

@@

\paragraph{Acknowledgement:}

This work was funded in part by National Science Foundation grant DMS 1007697.

>>>>

%\documentclass{article}
\documentclass[12pt]{article}
\usepackage{times}
%\usepackage{natbib}
%\usepackage{multicol}
\RequirePackage{natbib}
\usepackage{amsmath, amssymb, fullpage, amsthm, array, algorithm2e,graphicx,asa}
%\usepackage[dvips]{graphics}

@@

% Article top matter
\title{Validation of Visual Statistical Inference, Applied to Linear Models}
\author{
{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
\thanks{Mahbubul Majumder is Graduate Student, Heike Hofmann is Associate  Professor, Dianne Cook is Full Professor, Department of
Statistics and Statistical Laboratory, Iowa State University,
Ames, IA 50011-1210. This research is supported in part by the
National Science Foundation Grant \# DMS 1007697.}}
\date{\vspace{-.7in}}
%\date{\today}  %\today is replaced with the current date
\maketitle

\begin {abstract}  
Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis. Until recently, there were no formal visual methods in place for determining statistical significance of findings. This changed when \citet{buja:2009} conceptually introduced the lineup protocol for formal tests of visual findings. In this paper this is taken a step further by refining the terminology of visual inference, framing the protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists, and it is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models. A human subjects experiment is conducted using simulated data to provide controlled conditions. Results suggest that the lineup protocol provides results comparable to the conventional tests, out-performs them when data is contaminated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult trials.

{\bf Keywords: \sf visual inference, statistical graphics, lineup, non-parametric test} 
\end {abstract}

@@

\hh{Their minimum $p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}$ then follows a Beta distribution with shape parameters 1 and $m-1$, and corresponding distribution function 
\[
F_{p_0} (x) =  1- (1- x) ^{m-1}  \text{ for } x \in [0,1].
\]
\begin{eqnarray*}
P(p_D < p_{0}) &=& 1 - P(p_{0} \le p_D) = 1- \int_0^1  P(p_{0} \le p_D \mid p_D=t) f_{p_D}(t) dt   \\
&=& 1 - \int_0^1 F_{p_{0}}(t) f_{p_D}(t) dt = 1 - \int_0^1 f_{p_D}(t) dt + \int_0^1 (1-t)^{m-1} f_{p_D}(t) dt  \\
&=&  E\left[ (1 - p_D)^{m-1}\right].
\end{eqnarray*}

@@

%\paragraph{Acknowledgement:}
%This work was funded in part by National Science Foundation grant DMS 1007697.
